\documentclass[a4paper, 11pt,notoc]{article}
\pdfoutput=1
\usepackage{jcappub}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{verbatim}
\usepackage{caption}
\usepackage{xspace}
\usepackage{hyperref}

%\usepackage[skip=0cm,list=true,labelfont=it]{subcaption}
\usepackage{subcaption}
%\usepackage[sorting=none,backend=biber]{biblatex}



%\usepackage{draftwatermark}
%\SetWatermarkText{LHC DM WG DRAFT}
%\SetWatermarkLightness{0.9}
%\SetWatermarkScale{3.0}
\newcommand{\bra}[1]{\langle #1|}
\newcommand{\ket}[1]{|#1\rangle}
\newcommand{\sigv}{\ensuremath{\langle \sigma v_{\rm{rel}} \rangle}\xspace}
\newcommand{\MET}{\ensuremath{\slashed{E}_T}\xspace}
\newcommand{\mDM}{\ensuremath{m_{\rm{DM}}}\xspace}
\newcommand{\mmed}{\ensuremath{M_{\rm{med}}}\xspace}
\newcommand{\mMed}{\ensuremath{M_{\rm{med}}}\xspace}
\newcommand{\gDM}{\ensuremath{g_{\rm{DM}}}\xspace}
\newcommand{\gq}{\ensuremath{g_q}\xspace}
\newcommand{\gSM}{\gq}
\newcommand{\gdm}{\gDM}
\newcommand{\ifb}{\ensuremath{\rm{fb}^{-1}}\xspace}
\newcommand{\mA}{\ensuremath{M_{A}}\xspace}
\newcommand{\GamA}{\ensuremath{\Gamma_{A}}\xspace}
\newcommand{\Uli}{\color{red}}
\definecolor{cerulean}{RGB}{44,150,207}
\newcommand{\ATLASComments}{\color{cerulean}}
\newcommand{\dmsimp}{\textsc{DMsimp}\xspace}
\newcommand{\maddm}{\textsc{MadDM}\xspace}

%DIF PREAMBLE EXTENSION ADDED BY LATEXDIFF
%DIF UNDERLINE PREAMBLE %DIF PREAMBLE
\RequirePackage[normalem]{ulem} %DIF PREAMBLE
\RequirePackage{color}\definecolor{RED}{rgb}{1,0,0}\definecolor{BLUE}{rgb}{0,0,1} %DIF PREAMBLE
\providecommand{\DIFadd}[1]{{\protect\color{blue}\uwave{#1}}} %DIF PREAMBLE
\providecommand{\DIFdel}[1]{{\protect\color{red}\sout{#1}}}                      %DIF PREAMBLE
%DIF SAFE PREAMBLE %DIF PREAMBLE
\providecommand{\DIFaddbegin}{} %DIF PREAMBLE
\providecommand{\DIFaddend}{} %DIF PREAMBLE
\providecommand{\DIFdelbegin}{} %DIF PREAMBLE
\providecommand{\DIFdelend}{} %DIF PREAMBLE
%DIF FLOATSAFE PREAMBLE %DIF PREAMBLE
\providecommand{\DIFaddFL}[1]{\DIFadd{#1}} %DIF PREAMBLE
\providecommand{\DIFdelFL}[1]{\DIFdel{#1}} %DIF PREAMBLE
\providecommand{\DIFaddbeginFL}{} %DIF PREAMBLE
\providecommand{\DIFaddendFL}{} %DIF PREAMBLE
\providecommand{\DIFdelbeginFL}{} %DIF PREAMBLE
\providecommand{\DIFdelendFL}{} %DIF PREAMBLE
%DIF END PREAMBLE EXTENSION ADDED BY LATEXDIFF


\begin{document}
\title{\begin{boldmath} \huge Recommendations of the LHC Dark Matter Working Group: Comparing LHC searches for heavy mediators of dark matter production in visible and invisible decay channels\vspace{7mm} \end{boldmath}}

%%%%%%

\affiliation[*]{Editor}

\author[1,*]{Andreas~Albert,}
\affiliation[1]{III. Physikalisches Institut A, RWTH Aachen University, Aachen, Germany}

\author[2]{Mihailo Backovi\'c,}
\affiliation[2]{Center for Cosmology, Particle Physics and Phenomenology - CP3, Universite Catholique de Louvain, Louvain-la-neuve, Belgium}

\author[3,*]{Antonio~Boveia,}
\affiliation[3]{Ohio State University, 191 W. Woodruff Avenue
Columbus, OH 43210}
\emailAdd{antonio.boveia@cern.ch}

\author[4,*]{Oliver~Buchmueller,}
\affiliation[4]{High Energy Physics Group, Blackett Laboratory, Imperial College, Prince Consort Road, London, SW7 2AZ, United Kingdom}
\emailAdd{oliver.buchmueller@cern.ch}

\author[5,*]{Giorgio Busoni,} 
\affiliation[5]{ARC Centre of Excellence for Particle Physics at the Terascale, School of Physics, University of Melbourne, 3010, Australia}

\author[6,7]{Albert De~Roeck,}
\affiliation[6]{Antwerp University, BÐ2610 Wilrijk, Belgium. }
\affiliation[7]{CERN, EP Department, CH-1211 Geneva 23, Switzerland}

\author[8,*]{Caterina~Doglioni,}
\affiliation[8]{Fysiska institutionen, Lunds universitet, Lund, Sweden}
\emailAdd{caterina.doglioni@cern.ch}

\author[9,*]{Tristan~DuPree,}
\affiliation[9]{Nikhef, Science Park 105, NL-1098 XG Amsterdam, The Netherlands}


\author[10,*]{Malcolm Fairbairn,}
\affiliation[10]{Physics, King's College London, Strand, London, WC2R 2LS, UK}

\author[11]{\\ Marie-Helene Genest,}
\affiliation[11]{Laboratoire de Physique Subatomique et de Cosmologie, Université Grenoble-Alpes, \\ CNRS/IN2P3, 53 rue des Martyrs, 38026 Grenoble Cedex, France}

\author[12]{Stefania Gori,}
\affiliation[12]{Department of Physics, University of Cincinnati, Cincinnati, Ohio 45221, USA}

\author[13]{Giuliano~Gustavino,}
\affiliation[13]{Universita' di Roma Sapienza, Piazza Aldo Moro, 2, 00185 Roma, Italy e INFN}

\author[14,*]{Kristian~Hahn,}
\affiliation[14]{Department of Physics and Astronomy, Northwestern University, Evanston, Illinois 60208, USA}
\emailAdd{kristian.hahn@cern.ch}

\author[15,16,*]{Ulrich~Haisch,}
\affiliation[15]{Rudolf Peierls Centre for Theoretical Physics, University of Oxford, Oxford, OX1 3PN, United Kingdom}
\affiliation[16]{CERN, TH Department, CH-1211 Geneva 23, Switzerland}
\emailAdd{ulrich.haisch@physics.ox.ac.uk}

\author[7]{Philip~C.~Harris,} 

\author[17]{Dan~Hayden,}
\affiliation[17]{Michigan State University, 220 Trowbridge Rd, East Lansing, MI 48824,
USA}

\author[18]{Valerio~Ippolito,} 
\affiliation[18]{Laboratory for Particle Physics and Cosmology, Harvard University, USA}

\author[8]{Isabelle~John,}

\author[19,*]{Felix~Kahlhoefer,}
\affiliation[19]{DESY, Notkestra\ss e 85, D-22607 Hamburg, Germany}

\author[20]{Suchita~Kulkarni,}
\affiliation[20]{Institut f\"ur Hochenergiephysik, \"Osterreichische Akademie der Wissenschaften, \\ Nikolsdorfer Gasse 18, 1050 Wien, Austria}

\author[21]{Greg Landsberg,}
\affiliation[21]{Brown University, Dept. of Physics, 182 Hope St, Providence, RI 02912, USA.}

\author[22]{Steven Lowette,}
\affiliation[22]{Physics Department, Vrije Universieit Brussel, Brussels, Belgium}

\author[11]{Kentarou~Mawatari,}

\author[23]{Antonio Riotto,}
\affiliation[23]{Departement de Physique Theorique (DPT) Geneva Cosmology and Astroparticle Group and Centre for Astroparticle Physics  (CAP), 24 quai Ernest Ansermet
CH-1211 Geneve}

\author[24]{William~Shepherd,}
\affiliation[24]{Institut f\"ur Physik, Johannes-Gutenberg-Universit\"at Mainz, Staudingerweg 7, D-55128 Mainz, Germany}

\author[25,*]{Tim~M.P.~Tait,}
\affiliation[25]{Department of Physics and Astronomy, University of California, Irvine, California 92697, USA}
\emailAdd{ttait@uci.edu}

\author[3]{Emma Tolley,}

\author[10,*]{\\ Patrick Tunney,}

\author[26,*]{Bryan~Zaldivar,}
\affiliation[26]{Laboratoire de Physique Theorique d'Annecy-le-Vieux (LAPTH) 9 Chemin de Bellevue, B.P. 110, F-74941 Annecy-le-Vieux, CEDEX, France}

\author[24]{Markus~Zinser}


\hfill CERN-LPCC-2017-01


\abstract{
Weakly-coupled TeV-scale particles may mediate the interactions between normal matter and dark matter. If so, the LHC would produce dark matter through these mediators, leading to the familiar ``mono-$X$" search signatures, but the mediators would also produce signals without missing momentum via the same vertices involved in their production. This document from the LHC Dark Matter Working Group suggests how to compare searches for these two types of signals in case of vector and axial-vector mediators, based on a workshop that took place on September 19/20, 2016 and subsequent discussions. These suggestions include how to extend the spin-1 mediated simplified models already in widespread use to include lepton couplings. This document also provides analytic calculations of the relic density in the simplified models and reports an issue that arose when ATLAS and CMS first began to use preliminary numerical calculations of the dark matter relic density in these models.}
\maketitle


%%%%%%%%%%%%%%%%
\input{Introduction.tex}
%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%
\input{Models.tex}
%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%
\input{relic-11112016.tex}
%%%%%%%%%%%%%%%%

\section{Conclusions}

Simplified DM models in which the interactions between the visible and the invisible sector are mediated by the exchange of TeV-scale particles represent an interesting class of all  possible realizations of physics beyond the SM with a  viable DM candidate. In such scenarios the decays of the mediators can lead to final states where DM is  accompanied by SM radiation (so-called mono-$X$ events) but also to signatures that feature only SM particles (such as for instance di-jet or di-lepton events). To map out the DM parameter space in a given simplified model comparing and combining the different LHC search strategies is an important task. 

This document describes simplified spin-1 models where a $s$-channel mediator couple to DM, quarks, and leptons. Four benchmark scenarios with  different relative sizes of quark and lepton couplings are recommended that exemplify the rich phenomenology of the simplified DM model in the mono-jet, di-jet, and di-lepton channels. In this document, benchmark points for the vector and axial-vector mediator models where the mediator width is large have not been considered. Even though these may evade the constraints provided by di-jet and di-lepton resonance searches, LHC searches for non-resonant phenomena in the same final states, such as the ones in~\cite{ATLAS:2015nsi,CMS-PAS-EXO-15-009,Aad:2014wca, Khachatryan:2015scf}, are sensitive to wide mediators and can be subject of dedicated future studies.

This document also presents improved numerical calculations for the relic density for the dominant annihilation processes  that  involve  spin-0  and  spin-1  mediators. The full numerical results are compared to analytical calculations and found to be in good agreement in all cases  where the annihilation cross section is dominated by a single channel. It is furthermore shown that the omission of $t$-channel annihilation contributions can lead to parameter regions which feature an  erroneous  relic overabundance or  underabundance. 

 

\section{Acknowledgement} 
The work of CD is supported by the European Research Council under the European Union's Horizon 2020 research and innovation programme (ERC grant agreement No. 679305 DARKJETS). UH acknowledges the hospitality and support of the CERN theory division. MF and PT are funded by the European Research Council under the European Union's Horizon 2020 program (ERC grant agreement No. 648680 DARKHORIZONS). AR is supported by the Swiss National Science Foundation (SNSF), project title: `` Investigating the Nature of Dark Matter'', project number: 200020-159223. The work of TMPT is supported in part by National Science Foundation grants PHY-1316792 and PHY-1620638. The work of GL is partially supported by the DOE Award DE-SC0010010.
 

\bibliography{ref}
\bibliographystyle{JHEP}



\end{document}
