\section{Lepton couplings for simplified %Uli: dark matter 
DM models}
\label{sec:models}

%The model choices made for the early Run-2 LHC searches by the ATLAS/CMS DM Forum~\cite{Abercrombie:2015wmb} assume that the DM particle is a Dirac fermion~$\chi$.
%The new particle mediating the interaction (the ``mediator") is either a vector, an axial vector, a scalar or a pseudo-scalar, and it can be exchanged either in the $s$-channel or in the $t$-channel. In the limit of large $\mmed$, these (and all) models converge to a universal set of operators in an effective field theory (EFT)
%\cite{Beltran:2010ww,Goodman:2010yf,Bai:2010hh,Goodman:2010ku,Rajaraman:2011wf,Fox:2011pm,Bell:2015sza}.

% KH, reorganized the above


The simplified models recommended by the ATLAS/CMS DMF~\cite{Abercrombie:2015wmb} assume that DM is a Dirac fermion~$\chi$ and there is an additional heavy particle mediating the SM-DM interaction (the ``mediator"). In the most basic set of these models, the mediator is a vector, an axial-vector, a scalar or a pseudo-scalar boson. So far, ATLAS and CMS have focused on the subset of the models where the mediator is exchanged in the $s$-channel. These models contain four free parameters. In the vector and axial-vector models, the parameters are the DM mass~$\mDM$, the mediator mass~$\mmed$, the coupling~$\gDM$ of a mediator-DM-DM vertex, and the coupling~$\gq$ universal to all mediator-quark-quark vertices. In the scalar and pseudo-scalar models, a quark-mass-dependent Yukawa factor scales the coupling of the mediator-quark-quark vertices to avoid violating flavor constraints. These four quantities parameterize the production rate of the mediator in proton-proton collisions, its quark and DM decay rates, and the kinematic distributions of signal events.

Complete models of DM can contain mediators that may have (or require for consistency) couplings to other SM particles that are not found in the simplified models above. Such couplings would introduce additional decay modes of the mediator at the LHC as well as further DM annihilation channels in the relic density calculation. In this section, we discuss why and how to add lepton couplings to the vector and axial-vector simplified models, provide formulas for the total decay width of the mediators, and discuss the implementations of these models that are currently available. We then propose four benchmark scenarios for comparing di-jet, di-lepton, and mono-$X$ searches, based on rough estimates of the sensitivity of these searches with $30 \, {\rm fb}^{-1}$ of LHC data. We also comment on the interference between the mediator di-lepton process and the Drell-Yan backgrounds to di-lepton searches. 

%\footnote{An orthogonal set of models describe $t$-channel exchange \cite{Chang:2013oia,An:2013xka,Bai:2013iqa,DiFranzo:2013vra}. This class of simplified DM models is left for future iterations and will thus not be discussed in the following. \textcolor{red}{KH -- I think we will include this in the paper ...}
%\textcolor{cerulean}{CD -- yes, but I've removed for now since this will go to the dilepton analysers}} 

%CD: we have to discuss whether the following is included in this document
%, and mixing with the Higgs boson in scalar-mediated simplified models.  This recommendedation also briefly addresses t-channel (refs?) and two-Higgs doublet models (2HDM, refs?), which should be studied in the near term so that LHC searches can be interpreted in this expanded class of models after Moriond 2017.

\input{VandAV}
%\input{SMM}
%\input{2HDM}
%\input{Tchannel}


